### 2025-10-25 paper/header.tex
A bit of work on the setup

% **************GENERATED FILE, DO NOT EDIT**************

\title{Groups and smooth geometry using LieGroups.jl}

\author[1]{Ronny Bergmann}
\affil[1]{Department of Mathematical Sciences, Norwegian University of Science and Technology, Trondheim, Norway}

\keywords{Julia, Riemannian manifolds, Lie groups, Differential geometry}

\hypersetup{
pdftitle = {Groups and smooth geometry using LieGroups.jl},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {Ronny Bergmann},
pdfkeywords = {Julia, Riemannian manifolds, Lie groups, Differential geometry},
}